Wolbachia and Two Other Microbes Nobody Cares About That Changed The World Kinda

### 2015-12-16 paper.tex
Migrating to biblatex

@@

The bacteria is found in the ovaries and testes of a wide range of arthropods.[8] It is passed on via reproduction, and can only be spread through females. \textit{Wolbachia} are considered mutualistic endosymbionts, and are "required for survival of their hosts."[4] An interesting thing to note is how universal \textit{Wolbachia}seems to be. It has been demonstrated that\textit{Wolbachia} infects 25-70\% of species of insects.[7]

To give a bit of background,\textit{Wolbachia} was first discovered by Samuel Burt Wolbach and described by Marshall Hertig in 1924.[16] Wolbach had been studying the cause of Rocky Mountain Spotted Fever and by 1916 had shown that \textit{Dermacentor andersoni}, a tick native to the Rocky Mountain area, was the transmitter of the disease.[1] He was unable to grow the actual bacteria causing the disease, a species of\textit{Rickettsia}, in a cell-free culture leading "him to speculate as to the relationship between the cells of the host and the intracellular parasites."[2] This is important to note, because the discovery of\textit{Wolbachia} is closely linked to this idea.

Hertig came to work with Wolbach, and together they investigated a bacteria living in the gonads of the\textit{C. pipiens} mosquito. This bacteria is known today as\textit{Wolbachia pipientis} which Hertig gave a detailed description of in 1936.[3] At this point scientists did not know of the gene manipulation that\textit{Wolbachia} has on it's hosts, and the discovery largely went unnoticed.

@@

\section*{References}
[1] \_Wolbachia: A bug's Life in another Bug\_, by A. Hoerauf and R. U. Rao;  p.3

[2] \_Wolbachia: A bug's Life in another Bug\_, by A. Hoerauf and R. U. Rao;  p.3

[3] \_The Rickettsia, Wolbachia pipientis (gen. et sp.n.) and Associated Inclusions of the Mosquito, Culex pipiens\_, by Marshall Hertig

[4] \_The Discovery of Wolbachia in Arthorpods and Nematodes - A Historical Perspective\_, Wieslaw J. Kozek, Ramakrishna U. Rao  -- p1

[5] \_Wolbachia: A bug's Life in another Bug\_, by A. Hoerauf and R. U. Rao;  p.4

[6] Sironi M, Bandi C, Sacchi L, Di Sacco B, Damiani G, Genchi C: A close relative of the arthropod endosymbiont Wolbachia in a filarial worm. Mol Biochem Parasitol 1995; 74:223–227.

[7] \_The Discovery of Wolbachia in Arthorpods and Nematodes - A Historical Perspective\_, Wieslaw J. Kozek, Ramakrishna U. Rao  -- p1

[8] \_Biology of Wolbachia\_, John H. Werren  -- p1

[9] Clancy DJ, Hoffmann AA. 1996. Cytoplasmic incompatibility in Drosophila simulans: evolving complexity. Trends Ecol. Syst. 11:145–46

[10] O’Neill SL, Karr TL. 1990. Bidirectional incompatibility between conspecific populations ofDrosophila simulans. Nature 348:178–80

[11] Breeuwer JAJ, Stouthamer R, Burns DA, Pelletier DA, Weisburg WG, Werren JH. 1992. Phylogeny of cytoplasmic incompatibility microorganisms in the parasitoid wasp genus Nasonia (Hymenoptera: Pteromalidae) based on 16s ribosomal DNA sequences. Insect. Mol. Biol. 1:25–36

[12] Breeuwer JAJ, Werren JH. 1990. Microorganisms associated with chromosome destruction and reproductive isolation between two insect species. Nature 346:558–60

[13] Stouthamer R, Breeuwer JAJ, Luck RF, Werren JH. 1993. Molecular identification of microorganisms associated with parthenogenesis. Nature 361:66–68

[14] Rousset F, Bouchon D, Pintureau B, Juchault P, Solignac M. 1992. Wolbachia endosymbionts responsible for various alterations of sexuality in arthropods. Proc. R. Soc. London Ser. B 250:91–98

[15] Hurst G., Jiggins F. M., Graf von Der Schulenburg J. H., Bertrand D., et al. (1999). "Male killing Wolbachia in two species of insects". Proceedings of the Royal Society B 266 (1420): 735–740.

[16] Hertig M, Wolbach SB. 1924. Studies on rickettsia-like microorganisms in insects. J. Med. Res. 44:329–74

[17] Laven H. 1959. Speciation by cytoplasmic isolation in the Culex pipiens complex. Cold Spring Harbor Symp. Quant. Biol. 24:166–73

[18] Yen JH, Barr AR. 1971. New hypothesis of the cause of cytoplasmic incompatibiity in Culex pipiens . Nature 232:657–58

[19] Werren JH, Guo L, Windsor DW. 1995. Distribution of Wolbachia in neotropical arthropods. Proc. R. Soc. London Ser. B 262:147–204

[20] Werren JH, Biology of Wolbachia, Annual Rev. Entomol; 1997,42:587-609

[21] Werren JH, Zhang W, Guo LR. 1995. Evolution and phylogeny of Wolbachia : reproductive parasites of arthopods. Proc. R. Soc. London Ser. B 251:55–71

[22] Mercot H, Llorente B, Jacques M, Atlan A, Montchamp-Moreau C. 1995. Variability within the seychelles cytoplasmic incompatibility system in Drosophila simulans . Genetics 141:1015–23

[23] Werren JH, Biology of Wolbachia, Annual Rev. Entomol; 1997,42:587-609

[24] Werren JH, Biology of Wolbachia, Annual Rev. Entomol; 1997,42:587-609

[25] Stouthamer R, Breeuwer JAJ, Luck RF, Werren JH. 1993. Molecular identification of microorganisms associated with parthenogenesis. Nature 361:66–68

[26] Werren JH, Biology of Wolbachia, Annual Rev. Entomol; 1997,42:587-609

[27] Stouthamer R, Kazmer DJ. 1994. Cytogenetics of microbe-associated parthenogenesis and its consequences for gene flow in Trichogramma wasps. Heredity 73:317–27

>>>>

\usepackage[backend=biber]{biblatex}
\addbibresource{references.bib}

@@

The bacteria is found in the ovaries and testes of a wide range of arthropods.\cite{Wbio} It is passed on via reproduction, and can only be spread through females. \textit{Wolbachia} are considered mutualistic endosymbionts, and are "required for survival of their hosts."\cite{Wdisc_nem} An interesting thing to note is how universal \textit{Wolbachia}seems to be. It has been demonstrated that\textit{Wolbachia} infects 25-70\% of species of insects.\cite{Wdisc_nem}

To give a bit of background,\textit{Wolbachia} was first discovered by Samuel Burt Wolbach and described by Marshall Hertig in 1924.\cite{Winit} Wolbach had been studying the cause of Rocky Mountain Spotted Fever and by 1916 had shown that \textit{Dermacentor andersoni}, a tick native to the Rocky Mountain area, was the transmitter of the disease.\cite{wolbachia} He was unable to grow the actual bacteria causing the disease, a species of\textit{Rickettsia}, in a cell-free culture leading "him to speculate as to the relationship between the cells of the host and the intracellular parasites."\cite{wolbachia} This is important to note, because the discovery of\textit{Wolbachia} is closely linked to this idea.

Hertig came to work with Wolbach, and together they investigated a bacteria living in the gonads of the\textit{C. pipiens} mosquito. This bacteria is known today as\textit{Wolbachia pipientis} which Hertig gave a detailed description of in 1936.\cite{Wdiscription} At this point scientists did not know of the gene manipulation that\textit{Wolbachia} has on it's hosts, and the discovery largely went unnoticed.

@@

\section*{References}
\medskip
\printbibliography
